Prognostic and biological relevance of collagen-related genes in prostate cancer

### 2025-02-23 paper/markdown/manuscript_supplement.Rmd
Supplementary tables for survival modeling

flextable::flextable(suppl_tab$pool) %>% 
  set_widths(c(3.5, rep(3, 4), 2.3, 2.3)) %>% 
  footnote(1, 1, 
           value = as_paragraph('PSA: prostate-specific antigen; pT stage: pathological tumor stage; pN stage: pathological lymph node stage; pM stage: pathological metastasis stage; ISUP grade: grading according to the International Society of Urological Pathology.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 6:7, 
            value = as_paragraph("Categorical variables: \u03C7\u00B2 test with Cramer's V effect size statistic. Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. Biochemical relapse-free survival: Peto-Peto test. P values corrected for multiple testing with the false discovery rate method."), 
            part = 'header', 
            ref_symbol = 'b') %>% 
  theme_vanilla

@@

suppl_tab$nets %>% 
  group_by(Cohort) %>% 
  slice_max(`Hub score`, n = 5) %>% 
  ungroup %>% 
  flextable %>% 
  merge_v(1) %>% 
  italic(j = 2) %>% 
  set_widths(c(2.3, 3, 3.5, rep(2.3, 4))) %>% 
  footnote(1, 4, 
           value = as_paragraph("Degree: number of network neighbors, i.e. correlations with Spearman's \u03C1 >= 0.5."), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 5, 
           value = as_paragraph("Hub score: eigenvector of the affinity matrix; a metric of the overall correlation strength."), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  footnote(1, 6, 
           value = as_paragraph('Betweenness: minimum/maximum scaled number of the shortest paths between pairs of network vertices which pass through the vertex of interest; measure of connectivity and centrality.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  footnote(1, 7, 
           value = as_paragraph('Transitivity: probability that the adjacent vertices of a vertex are connected; measure of the local densitiy of the co-expression network.'), 
           part = 'header', 
           ref_symbols = 'd') %>% 
  theme_vanilla

@@

>>>>

flextable::flextable(suppl_tab$pool) %>% 
  set_widths(c(3.5, rep(3.3, 4), 2.2, 2.2)) %>% 
  footnote(1, 1, 
           value = as_paragraph('PSA: prostate-specific antigen; pT stage: pathological tumor stage; pN stage: pathological lymph node stage; pM stage: pathological metastasis stage; ISUP grade: grading according to the International Society of Urological Pathology.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 6:7, 
            value = as_paragraph("Categorical variables: \u03C7\u00B2 test with Cramer's V effect size statistic. Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. Biochemical relapse-free survival: Peto-Peto test. P values corrected for multiple testing with the false discovery rate method."), 
            part = 'header', 
            ref_symbol = 'b') %>% 
  theme_vanilla

@@

suppl_tab$nets %>% 
  group_by(Cohort) %>% 
  slice_max(`Hub score`, n = 5) %>% 
  ungroup %>% 
  flextable %>% 
  merge_v(1) %>% 
  italic(j = 2) %>% 
  set_widths(c(2, 2.7, 3.5, rep(2.5, 4))) %>% 
  footnote(1, 4, 
           value = as_paragraph("Degree: number of network neighbors, i.e. correlations with Spearman's \u03C1 >= 0.5."), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 5, 
           value = as_paragraph("Hub score: eigenvector of the affinity matrix; a metric of the overall correlation strength."), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  footnote(1, 6, 
           value = as_paragraph('Betweenness: minimum/maximum scaled number of the shortest paths between pairs of network vertices which pass through the vertex of interest; measure of connectivity and centrality.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  footnote(1, 7, 
           value = as_paragraph('Transitivity: probability that the adjacent vertices of a vertex are connected; measure of the local densitiy of the co-expression network.'), 
           part = 'header', 
           ref_symbols = 'd') %>% 
  theme_vanilla

@@

```{r tab-uni-bcr, tab.cap = "Univariable analysis of biochemical relapse (BCR) free survival for the collagen-realted transcripts in the pooled GEO, TCGA, and DKFZ cohort. PCA patients were classified as high and low expressors of the collagen-related transcripts by expression cutoffs corresponding to the largest differences in BCR-free survival between the expression strata. BCR risk of high as compared with low expressors was modeled by univariable Cox proportional hazard regression. P values were corrected for multiple testing with the false discovery rate method. The modeling results are shown for the collagen-related genes associated with significant differences in BCR-free survival in all three investigated cohorts. The full analysis is available as a supplementary Excel file."}

suppl_tab$uni_bcr %>% 
  filter(`Gene symbol` %in% reduce(uni_cut$common_significant, union)) %>% 
  arrange(`Association with BCR risk`, `Gene symbol`, Cohort) %>% 
  relocate(`Association with BCR risk`, `Gene symbol`, Cohort) %>% 
  flextable %>% 
  merge_v(1:2) %>% 
  italic(j = 2) %>% 
  set_widths(c(2, 2, 2, 1.8, rep(2.4, 2), 2.2, 2.3, rep(1.7, 2))) %>% 
  footnote(1, 7, 
           value = as_paragraph('Harazd ratio (HR) with 95% confidence intervals (CI).'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 9, 
           value = as_paragraph("C-index: Harrel's concordance index, a measure of model accuracy. High values are expected for a highly accurate model; concordance index = 0.5 is expected for random predictions."), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  footnote(1, 10, 
           value = as_paragraph('IBS: Integrated Brier Score, a metric of overall model calibration. Low values are expected for a well calibrate model. Intergrated Brier score = 0.25 is expected for random predictions.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  theme_vanilla

```{r tab-tune, tab.cap = "Selection of the optimal parameters of machine learning models of biochemical relapse (BCR) free survival in the pooled GEO training cohort."}

flextable::flextable(suppl_tab$tune) %>% 
  set_widths(c(2.7, 6, 6)) %>% 
  footnote(1, 1, 
           value = as_paragraph('SVM: Support Vector Machines; GBM: Gradient Boosted Machines.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1:6, 1, 
           value = as_paragraph('Explanatory factors: expression levels of the collagen-related transcripts.'), 
           part = 'body', 
           ref_symbols = 'b') %>% 
  footnote(7, 1, 
           value = as_paragraph('Explanatory factors: pathological tumor stage, ISUP grade, and log2 expression of the PSA coding gene KLK3 in the tumor tissue.'), 
           part = 'body', 
           ref_symbols = 'c') %>% 
  footnote(8, 1, 
           value = as_paragraph('Explanatory factors: pathological tumor stage, ISUP grade, and log2 expression of the PSA coding gene KLK3 in the tumor tissue, log2 expression of the collagen-related transcripts.'), 
           part = 'body', 
           ref_symbols = 'd') %>% 
  theme_vanilla

```{r tab-bcr-eval, tab.cap = "Performance of machine learning models at prediction of biochemical relapse (BCR) free survival."}

flextable::flextable(suppl_tab$bcr_eval) %>% 
  merge_v(1:2) %>% 
  set_widths(c(3.5, 2.7, 3.5, rep(2.5, 2))) %>% 
   footnote(1, 1, 
           value = as_paragraph('SVM: Support Vector Machines; GBM: Gradient Boosted Machines.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1:18, 1, 
           value = as_paragraph('Explanatory factors: expression levels of the collagen-related transcripts.'), 
           part = 'body', 
           ref_symbols = 'b') %>% 
  footnote(19:21, 1, 
           value = as_paragraph('Explanatory factors: pathological tumor stage, ISUP grade, and log2 expression of the PSA coding gene KLK3 in the tumor tissue.'), 
           part = 'body', 
           ref_symbols = 'c') %>% 
   footnote(22:24, 1, 
           value = as_paragraph('Explanatory factors: pathological tumor stage, ISUP grade, and log2 expression of the PSA coding gene KLK3 in the tumor tissue, log2 expression of the collagen-related transcripts.'), 
           part = 'body', 
           ref_symbols = 'd') %>% 
  footnote(1, 4, 
           value = as_paragraph("C-index: Harrel's concordance index, a measure of model accuracy. High values are expected for a highly accurate model; concordance index = 0.5 is expected for random predictions."), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  footnote(1, 5, 
           value = as_paragraph('IBS: Integrated Brier Score, a metric of overall model calibration. Low values are expected for a well calibrate model. Intergrated Brier score = 0.25 is expected for random predictions.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  theme_vanilla

```{r tab-os-tune, tab.cap = "Selection of the optimal parameters of machine learning models of overall survival (OS) in the training subset of the GSE16560 cohort. The models used log2 expression of the collagen-related genes as explanatory variables."}

flextable::flextable(suppl_tab$os_tune) %>% 
  set_widths(c(2.7, 6, 6)) %>% 
  footnote(1, 1, 
           value = as_paragraph('SVM: Support Vector Machines; GBM: Gradient Boosted Machines.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  theme_vanilla

```{r tab-os-eval, tab.cap = "Performance of machine learning models at prediction of overall survival in the training and test subsets of the GSE16560 cohort. The models used log2 expression of the collagen-related genes as explanatory factors."}

flextable::flextable(suppl_tab$os_eval) %>% 
  set_widths(c(3.5, 2.7, rep(2.5, 2))) %>% 
  merge_v(1) %>% 
  footnote(1, 3, 
           value = as_paragraph("C-index: Harrel's concordance index, a measure of model accuracy. High values are expected for a highly accurate model; concordance index = 0.5 is expected for random predictions."), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  footnote(1, 4, 
           value = as_paragraph('IBS: Integrated Brier Score, a metric of overall model calibration. Low values are expected for a well calibrate model. Intergrated Brier score = 0.25 is expected for random predictions.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  theme_vanilla
